### 2025-03-04 Exam Assignments/Chapters/02/02.tex
Week 09
Refining all chapters

    ~\\
    In the same way, false sharing occurs in computing when multiple threads modify different parts of the same cache line. Even though each thread may be working on separate variables, they're forced to constantly invalidate and reload the cache line they "share," leading to substantial inefficiencies.

    ~\\
    Here is an illustration of the sceario:

>>>>

    ~\\
    In the same way, false sharing occurs in computing when multiple threads modify different parts of the same cache line. Even though each thread may be working on separate variables, they're forced to constantly invalidate and reload the cache line they "share", leading to substantial inefficiencies.

    ~\\
    Here is an illustration of the scenario:

### 2025-03-04 Exam Assignments/Chapters/03/03.tex
Refining all chapters

~\\
In this code, a team of thread get created my the \em \#pragma omp parallel \em clause.
Their task is to print their respective thread numbers.
But since no order is defined, the order of their execution are "random", i.e. determined at runtime by the operating system.

@@

~\\
While still not defining the order of the tasks, they yet seem to execute in order.
Interestingly, this is not quite true, as the main reason the seem to do so is because the \em terminate \em in order.
This is simply due to the fact that the work load for a thread with a higher thread number is substantially higher than for another one with a lower number.
This is because fib(n+t) is invoked, but it takes very long to calculate because of the big n, and even a singe step by t causes fib(n+t) to take nearly twice as long.

\subsubsection{Game of Life}
The code in Listing 3.10 looks well designed.
It collapses the two nested for loops of size $size$ into a single one of size $size^2$.
Since each iteration approximately takes the same amount of work, the implicit static scheduling ensures a good distribution of the workload.
This way, every thread has to compute \( \frac{size^2}{\text{omp\_get\_num\_threads()}} \) iterations.

@@

~\\
\subsubsection{Optimizing Random Shooting}
When trying to approximate $\pi$ by the method of random shooting we use random numbers.
They, however, don't need to be truly random, they just have to simulate it well enough.
One way to do this simulation of randomness is the rnd function like it is defined in the book (or see the code below).
It calculates a random number quite efficiently.
Another benfit is that we do not need multiple instances of random number generators, and we also don't run the risk of misusing those by calling their methods in different threads.
The code is shown in the \hyperref[fig:calculate_pi_v2]{figure} below.
It runs very quickly and calculates $\pi$ very precisely.
The running time on my particular machine was 0.0313137 s, in comparison to the previous version with a running time of 0.117771 s.

>>>>

~\\
In this code, a team of threads gets created by the \em \#pragma omp parallel \em clause.
Their task is to print their respective thread numbers.
But since no order is defined, the order of their execution are "random", i.e. determined at runtime by the operating system.

@@

~\\
While still not defining the order of the tasks, they yet seem to execute in order.
Interestingly, this is not quite true, as the main reason they seem to do so is because their \em termination \em is in order.
This is simply due to the fact that the work load for a thread with a higher thread number is substantially higher than for another one with a lower number.
This is because fib(n+t) is invoked, but it takes very long to calculate because of the big n, and even a singe step by t causes fib(n+t) to take nearly twice as long.

~\\
\subsubsection{Game of Life}
The code in Listing 3.10 looks well designed.
It collapses the two nested for loops of size $size$ into a single one of size $size^2$.
Since each iteration approximately takes the same amount of work, the implicit static scheduling ensures a good distribution of the workload.
This way, every thread has to compute \( \frac{size^2}{\text{omp\_get\_num\_threads()}} \) iterations.

@@

~\\
\subsubsection{Optimizing Random Shooting}
When trying to approximate $\pi$ by the method of random shooting we use random numbers.
They, however, don't need to be truly random, they just have to simulate it well enough.
One way to do this simulation of randomness is the rnd function like it is defined in the book (or see the code below).
It calculates a random number quite efficiently.
Another benefit is that we do not need multiple instances of random number generators, and we also don't run the risk of misusing those by calling their methods in different threads.
The code is shown in the \hyperref[fig:calculate_pi_v2]{figure} below.
It runs very quickly and calculates $\pi$ very precisely.
The running time on my particular machine was 0.0313137 s, in comparison to the previous version with a running time of 0.117771 s.

### 2025-03-04 Exam Assignments/Chapters/04/04.tex
Refining all chapters

~\\
And here are the performance gains compared to std::sort():

@@

~\\
% Let us now analyze the second two pictures, showing the running times at different array sizes.
% The graphs are as expected, the running times of the parallel algorithms is better, even more so for very large array sizes, as in those instances we can fully employ the parallel power of the cpu cores, thus decreasing the running time significantly.
% This is why the speedup gets better with increasing array size.
% One might think that this speedup is exponential, but it is not (note the logarithmic scale!), it is rather approximately $(\text{Serial Running Time}) \cdot (1 - \frac{1}{12 \cdot E})$, where $E$ is the efficiency $E = \frac{\frac{\text{Serial Running Time}}{\text{Threads}}}{\text{Parallel Running Time}}$, and note that the number of threads on this machine is 12.
% Note, that Serial Running Time is expected to grow $\Theta(n \cdot \log(n))$, and so should the speedup.
Let us now analyze the second two pictures, showing the running times at different array sizes.
The graphs are as expected, the running times of the parallel algorithms is better, even more so for very large array sizes, as in those instances we can fully employ the parallel power of the cpu cores without much overhead relative to the running time, thus decreasing the running time significantly.
This is why the speedup gets better with increasing array size initially.
The speedup, however, start to plateau quickly at an array size of roughly 1 MB, as at this point the overhead of parallelization with respect to the running time stays constant (i.e. $\frac{\text{Time of Overhead}}{\text{Parallel Running Time}} = \text{const.}$):

@@

\newpage
\subsection{What every systems programmer should know about concurrency}
\subsubsection{Why it is not enough to declare shared variables as volatile}
By declaring variables as volatile in C and C++, we make the compiler read and write the variable from and to memory every time we access it.
Cache coherency protocols then ensure that every local cache of the cpu cores has the same value of the shared variable.
So everything should be good then, right?

>>>>

~\\
And here are the performance gains compared to std::sort(). Note the different coloring:

@@

~\\
% Let us now analyze the second two pictures, showing the running times at different array sizes.
% The graphs are as expected, the running times of the parallel algorithms is better, even more so for very large array sizes, as in those instances we can fully employ the parallel power of the cpu cores, thus decreasing the running time significantly.
% This is why the speedup gets better with increasing array size.
% One might think that this speedup is exponential, but it is not (note the logarithmic scale!), it is rather approximately $(\text{Serial Running Time}) \cdot (1 - \frac{1}{12 \cdot E})$, where $E$ is the efficiency $E = \frac{\frac{\text{Serial Running Time}}{\text{Threads}}}{\text{Parallel Running Time}}$, and note that the number of threads on this machine is 12.
% Note, that Serial Running Time is expected to grow $\Theta(n \cdot \log(n))$, and so should the speedup.
Let us now analyze the second two pictures, showing the running times at different array sizes.
The graphs are as expected, the running times of the parallel algorithms is better, even more so for very large array sizes, as in those instances we can fully employ the parallel power of the CPU cores without much overhead relative to the running time, thus decreasing the running time significantly.
This is why the speedup gets better with increasing array size initially.
The speedup, however, start to plateau quickly at an array size of roughly 1 MB, as at this point the overhead of parallelization with respect to the running time stays constant (i.e. $\frac{\text{Time of Overhead}}{\text{Parallel Running Time}} = \text{const.}$):

@@

\newpage
\subsection{What every systems programmer should know about concurrency}
\subsubsection{Why it is not enough to declare shared variables as volatile}
By declaring variables as volatile in C and C++, we make the compiler read and write the variable from and to memory every time we access it.
Cache coherency protocols then ensure that every local cache of the CPU cores has the same value of the shared variable.
So everything should be good then, right?

### 2025-03-04 Exam Assignments/Chapters/08/08.tex
Refining all chapters

\begin{document}
    \section{Vector Intrinsics and Instruction-Level Parallelism}
    \subsection{Analyzing slide 14}
    Slide 14 explains how we can utilize vector instrinsics to transpose 8x8 matrices. Here is the code:
    \begin{lstlisting}
// transpose 8 x 8 matrix with intrinsics
inline void transpose_8x8_int_intrinsics(__m256i *vecs) {
    __m256 *v = reinterpret_cast<__m256 *>(vecs);
    __m256 a = _mm256_unpacklo_ps(v[0], v[1]);
    __m256 b = _mm256_unpackhi_ps(v[0], v[1]);
    __m256 c = _mm256_unpacklo_ps(v[2], v[3]);
    __m256 d = _mm256_unpackhi_ps(v[2], v[3]);
    __m256 e = _mm256_unpacklo_ps(v[4], v[5]);
    __m256 f = _mm256_unpackhi_ps(v[4], v[5]);
    __m256 g = _mm256_unpacklo_ps(v[6], v[7]);
    __m256 h = _mm256_unpackhi_ps(v[6], v[7]);
    auto tmp = _mm256_shuffle_ps(a, c, 0x4E);
    a = _mm256_blend_ps(a, tmp, 0xCC);
    c = _mm256_blend_ps(c, tmp, 0x33);
    tmp = _mm256_shuffle_ps(b, d, 0x4E);
    b = _mm256_blend_ps(b, tmp, 0xCC);
    d = _mm256_blend_ps(d, tmp, 0x33);
    tmp = _mm256_shuffle_ps(e, g, 0x4E);
    e = _mm256_blend_ps(e, tmp, 0xCC);
    g = _mm256_blend_ps(g, tmp, 0x33);
    tmp = _mm256_shuffle_ps(f, h, 0x4E);
    f = _mm256_blend_ps(f, tmp, 0xCC);
    h = _mm256_blend_ps(h, tmp, 0x33);
    v[0] = _mm256_permute2f128_ps(a, e, 0x20);
    v[1] = _mm256_permute2f128_ps(c, g, 0x20);
    v[2] = _mm256_permute2f128_ps(b, f, 0x20);
    v[3] = _mm256_permute2f128_ps(d, h, 0x20);
    v[4] = _mm256_permute2f128_ps(a, e, 0x31);
    v[5] = _mm256_permute2f128_ps(c, g, 0x31);
    v[6] = _mm256_permute2f128_ps(b, f, 0x31);
    v[7] = _mm256_permute2f128_ps(d, h, 0x31);
}
    \end{lstlisting}
Let us analyze how we can calculate a specific column, say $v[0]$. It depends on the variables $a$ and $e$, so let us track them:

@@

~\\
From here, we can see that we are almost done, $a$ and $e$ look like they almost contain one column, in fact the contain two columns "cut in the middle". Hence, after a final permutation we should be done calculating a column. To this end, we use \texttt{\_mm256\_permute2f128\_ps} and specify that we want to use the lower bits of both $a$ and $e$ by setting the mask to \texttt{CC}:
\begin{lstlisting}
    v[0] = _mm256_permute2f128_ps(a, e, 0x20);
    v[0] = {a00, a10, a20, a30, a40, a50, a60, a70};
\end{lstlisting}

@@

~\\
If one writes dependent code, i.e. the next line of code may only be executed after the current one because of a dependency, the latency of the operation will be the main factor for performance, since we want our result as quickly as possible to procede with the execution.

~\\
On the other hand, if we write independent code which also uses many repeated operations, throughput will be key, as it determines how many clock cycles we approximately need to finish all calculations.
%  \\ ($no\_clock\_cycles \approx \frac{no\_operations}{throughput\_of\_operation}$ (Note that the number of operations does not directly correlate with the number of instructions in code, as certain code instructions need more cpu instructions)).
In this case, the latency has a neglectable influence on the performance, especially if the number of operations grows bigger.

@@

\begin{description}
    \item[Vectorized Linear Probing] The keys should be integer values, so that they can fit comfortably into the vector registers. Linear probing starts by checking the index specified by the hash value of the current key first, and then keeps going one step at a time until we found our key. The "key" idea is to simply load multiple consecutive keys into one vector register, and to compare them all against the search key. If one comparison is successful, we can calculate the index and return tha value at that index. Otherwise, we keep probing with the next set of keys.

@@

    \item[Vectorized Fingerprinting] As mentioned, vectorized linear probing require integer keys. So what could we do if wanted to use strings instead? The "key" idea is to calculate "keys" for the keys themself (quite many keys around here). One can think of these keys as hash values (though not necessarily the same as the real hashes) of the keys. To avoid confusion, they are called \em fingerprints \em instead, and are quite short, in the paper the authors used fingerprints of size 8 or 16 bits. Using this technique, we can use vectorized linear probing on these fingerprints as described earlier, with the only difference being that we might run into fingerprint-conflicts. Thus, after a successful match of fingerprints, we have to then compare the actual keys to determine an actual hit.
\end{description}
\end{document}

>>>>

\begin{document}
    \section{Vector Intrinsics and Instruction-Level Parallelism}
    \subsection{Analyzing slide 14}
    Slide 14 explains how we can utilize vector intrinsics to transpose 8x8 matrices. Here is the code:
    \begin{lstlisting}
// transpose 8 x 8 matrix with intrinsics
inline void transpose_8x8_int_intrinsics(__m256i *vecs) {
    __m256 *v = reinterpret_cast<__m256 *>(vecs);
    __m256 a = _mm256_unpacklo_ps(v[0], v[1]);
    __m256 b = _mm256_unpackhi_ps(v[0], v[1]);
    __m256 c = _mm256_unpacklo_ps(v[2], v[3]);
    __m256 d = _mm256_unpackhi_ps(v[2], v[3]);
    __m256 e = _mm256_unpacklo_ps(v[4], v[5]);
    __m256 f = _mm256_unpackhi_ps(v[4], v[5]);
    __m256 g = _mm256_unpacklo_ps(v[6], v[7]);
    __m256 h = _mm256_unpackhi_ps(v[6], v[7]);
    auto tmp = _mm256_shuffle_ps(a, c, 0x4E);
    a = _mm256_blend_ps(a, tmp, 0xCC);
    c = _mm256_blend_ps(c, tmp, 0x33);
    tmp = _mm256_shuffle_ps(b, d, 0x4E);
    b = _mm256_blend_ps(b, tmp, 0xCC);
    d = _mm256_blend_ps(d, tmp, 0x33);
    tmp = _mm256_shuffle_ps(e, g, 0x4E);
    e = _mm256_blend_ps(e, tmp, 0xCC);
    g = _mm256_blend_ps(g, tmp, 0x33);
    tmp = _mm256_shuffle_ps(f, h, 0x4E);
    f = _mm256_blend_ps(f, tmp, 0xCC);
    h = _mm256_blend_ps(h, tmp, 0x33);
    v[0] = _mm256_permute2f128_ps(a, e, 0x20);
    v[1] = _mm256_permute2f128_ps(c, g, 0x20);
    v[2] = _mm256_permute2f128_ps(b, f, 0x20);
    v[3] = _mm256_permute2f128_ps(d, h, 0x20);
    v[4] = _mm256_permute2f128_ps(a, e, 0x31);
    v[5] = _mm256_permute2f128_ps(c, g, 0x31);
    v[6] = _mm256_permute2f128_ps(b, f, 0x31);
    v[7] = _mm256_permute2f128_ps(d, h, 0x31);
}
    \end{lstlisting}
Let us analyze how we can calculate a specific column, say $v[0]$. It depends on the variables $a$ and $e$, so let us track them:

@@

~\\
From here, we can see that we are almost done, $a$ and $e$ look like they almost contain one column, in fact the contain two columns "cut in the middle". Hence, after a final permutation we should be done calculating a column. To this end, we use \texttt{\_mm256\_permute2f128\_ps} and specify that we want to use the lower bits of both $a$ and $e$ by setting the mask to \texttt{0x20}:
\begin{lstlisting}
    v[0] = _mm256_permute2f128_ps(a, e, 0x20);
    v[0] = {a00, a10, a20, a30, a40, a50, a60, a70};
\end{lstlisting}

@@

~\\
If one writes dependent code, i.e. the next line of code may only be executed after the current one because of a dependency, the latency of the operation will be the main factor for performance, since we want our result as quickly as possible to proceed with the execution.

~\\
On the other hand, if we write independent code which also uses many repeated operations, throughput will be key, as it determines how many clock cycles we approximately need to finish all calculations.
%  \\ ($no\_clock\_cycles \approx \frac{no\_operations}{throughput\_of\_operation}$ (Note that the number of operations does not directly correlate with the number of instructions in code, as certain code instructions need more cpu instructions)).
In this case, the latency has a negligible influence on the performance, especially if the number of operations grows bigger.

@@

\begin{description}
    \item[Vectorized Linear Probing] The keys should be integer values, so that they can fit comfortably into the vector registers. Linear probing starts by checking the index specified by the hash value of the current key first, and then keeps going one step at a time until we found our key. The "key" idea is to simply load multiple consecutive keys into one vector register, and to compare them all against the search key. If one comparison is successful, we can calculate the index and return the value at that index. Otherwise, we keep probing with the next set of keys.

@@

    \item[Vectorized Fingerprinting] As mentioned, vectorized linear probing requires integer keys. So what could we do if wanted to use strings instead? The "key" idea is to calculate "keys" for the keys themselves (quite many keys around here). One can think of these keys as hash values (though not necessarily the same as the real hashes) of the keys. To avoid confusion, they are called \em fingerprints \em instead, and are quite short, in the paper the authors used fingerprints of size 8 or 16 bits. Using this technique, we can use vectorized linear probing on these fingerprints as described earlier, with the only difference being that we might run into fingerprint-conflicts. Thus, after a successful match of fingerprints, we have to then compare the actual keys to determine an actual hit.
\end{description}
\end{document}
